\section{Implementation}

\subsection{System Startup and Resource Allocation}
The entry point of the application is intentionally lightweight. In
\texttt{main.cpp}, the \texttt{setup()} function performs four steps:
\begin{enumerate}
  \item initialize the serial port and I\textsuperscript{2}C bus,
  \item allocate the application context,
  \item create queues, mutexes, and the Internet-ready semaphore, and
  \item spawn all RTOS tasks.
\end{enumerate}

This startup strategy is important because it keeps initialization
deterministic. The \texttt{loop()} function does not perform application logic;
it simply yields with a long delay. In other words, after boot the entire
system behavior is task-driven. This is a clear and appropriate RTOS pattern
for the given assignment.

The application context is initialized by \texttt{init\_app\_shared\_state()},
which sets default values such as invalid sensor readings, disconnected Wi-Fi
state, and the default LCD content mode. This avoids undefined startup
behavior.

\subsection{Shared State Helpers}
The file \texttt{global.cpp} implements helper functions to read and update the
compact shared state block. The core logic is simple but useful:
\begin{itemize}
  \item \texttt{lockSharedState()} and \texttt{unlockSharedState()} hide mutex
  handling details,
  \item \texttt{app\_set\_latest\_sensor()} updates the latest temperature and
  humidity,
  \item \texttt{app\_set\_wifi\_connected()} and
  \texttt{app\_get\_wifi\_connected()} manage connectivity state, and
  \item \texttt{app\_set\_lcd\_content\_mode()} and
  \texttt{app\_get\_lcd\_content\_mode()} coordinate LCD display policy.
\end{itemize}

The value of this module is not its complexity but its role in improving code
quality. It makes synchronization explicit and reduces the need for broad
\texttt{extern} globals. This is a meaningful improvement over directly
exposing mutable state to every file.

\subsection{Sensor Task}
The sensor task is the timing anchor of the whole application. It performs
periodic DHT20 acquisition every 2 seconds using \texttt{vTaskDelayUntil()}.
This API is a good design choice because it maintains a stable period relative
to the previous wake time, instead of accumulating drift through repeated
\texttt{delay()} calls.

The execution flow is as follows:
\begin{enumerate}
  \item initialize the DHT20 once, under I\textsuperscript{2}C mutex
  protection;
  \item on each period, attempt to lock the I\textsuperscript{2}C bus for up to
  300 ms;
  \item read temperature and humidity;
  \item if the data are valid, update the shared state and dispatch the sample
  to all consumers;
  \item if the data are invalid, propagate an invalid sample to the LCD and web
  server for graceful failure handling.
\end{enumerate}

One of the most elegant aspects of this task is the fan-out pattern. A single
physical read is turned into multiple queue writes:
\texttt{lcdQueue}, \texttt{webQueue}, \texttt{coreQueue}, and
\texttt{tinyMLQueue}. This means the sensor is sampled exactly once per period,
while the rest of the system remains modular.

\subsection{LED Manager Task}
The discrete LED task is implemented as an event-driven state machine. It does
not read sensors directly. Instead, it receives \texttt{led\_command\_t}
messages through \texttt{ledQueue}. The LED task supports two categories of
behavior:
\begin{itemize}
  \item continuous blinking according to the current temperature mode, and
  \item short pulse sequences for exceptional events.
\end{itemize}

The temperature-to-blink mapping is implemented through
\texttt{classify\_temperature\_mode()} and \texttt{led\_half\_period\_ticks()}:
\begin{itemize}
  \item cold: 5-second half-period,
  \item normal: 1-second half-period,
  \item hot: 100-ms half-period,
  \item error: 120-ms half-period.
\end{itemize}

The task also supports quick pulse bursts:
\begin{itemize}
  \item Wi-Fi connected: 3 fast blinks,
  \item error on: 6 fast blinks,
  \item error clear: stop pulsing and return to off.
\end{itemize}

The strong point of this design is that the LED module is both periodic and
reactive. It uses queue timeouts as implicit timing intervals, so it can either
wait for a new command or toggle according to the current mode without another
timer task.

\subsection{NeoPixel Task}
The NeoPixel task is simpler but still well designed. It blocks on
\texttt{neoQueue} and updates the RGB LED whenever a new humidity command
arrives. The function \texttt{humidity\_to\_color()} maps humidity into a
continuous gradient:
\begin{itemize}
  \item \(0\%\) to \(50\%\): red to green,
  \item \(51\%\) to \(100\%\): green to blue.
\end{itemize}

This is an improvement over a strict three-bin implementation because it
provides richer visual information while still satisfying the assignment's
three-level requirement. The task also supports a disable command that clears
the LED immediately, which is useful for future extension or power-saving
scenarios.

\subsection{LCD Task}
The LCD task owns the local user interface. It maintains the most recent sensor
sample and the most recent TinyML result, then decides what to show based on
system state. The logic has three levels:
\begin{enumerate}
  \item in AP mode, always show the sensor screen;
  \item in STA mode with normal LCD content mode, show the sensor screen;
  \item in STA mode with forecast LCD content mode, show the TinyML forecast.
\end{enumerate}

This conditional behavior is an excellent example of policy separation. The
task itself decides \textit{how} to render each screen, while the rest of the
system decides \textit{which content mode} is currently active.

The LCD module also reacts to environmental context:
\begin{itemize}
  \item if the sensor data are invalid, it prints \texttt{Sensor Error};
  \item if Wi-Fi is connected, it can display signal quality in percent and a
  qualitative label such as \texttt{GOOD} or \texttt{WEAK};
  \item if TinyML results exist, it shows forecast label and rain probability.
\end{itemize}

An important implementation detail is that LCD rendering is not performed at a
fixed blind period. Instead, the task updates when a new sensor sample arrives,
when TinyML changes, when RSSI changes meaningfully, or when AP/STA mode
changes. This avoids unnecessary I\textsuperscript{2}C traffic and makes the
display more event-aware.

\subsection{Web Server Task}
The web module is the largest subsystem in the repository. It combines
\texttt{WebServer}, \texttt{Preferences}, local runtime state, and Wi-Fi mode
management. Its implementation can be divided into five functional groups.

\subsubsection*{1) AP/STA workflow.}
The board starts either from the most recently saved Wi-Fi profile or directly
in AP mode if no profile exists. The helper functions
\texttt{startApMode()} and \texttt{startStaConnect()} reset Wi-Fi state,
configure the correct mode, and update internal status flags. If STA connection
fails after 10 seconds, the system automatically returns to AP mode. The BOOT
button also forces AP mode manually.

\subsubsection*{2) Persistent Wi-Fi profiles.}
The project uses \texttt{Preferences} to store up to five Wi-Fi profiles. The
most recently used profile is kept at index 0. The code supports:
\begin{itemize}
  \item reading saved profiles,
  \item updating or inserting a profile,
  \item removing all saved profiles.
\end{itemize}
This is a practical improvement because it lets the board remember STA
credentials across restarts without hard-coding a single network.

\subsubsection*{3) AP dashboard and settings pages.}
The root page in AP mode displays:
\begin{itemize}
  \item temperature and humidity,
  \item TinyML label and probability,
  \item AP IP and Wi-Fi status,
  \item target STA SSID and STA IP,
  \item external LED control,
  \item chart history, and
  \item quick Wi-Fi scan results.
\end{itemize}
The settings page expands the Wi-Fi management workflow by providing scan,
saved-profile, and connect forms. The STA page is intentionally simpler and
acts mainly as a status page rather than a full dashboard.

\subsubsection*{4) REST endpoints.}
The task registers multiple routes:
\begin{itemize}
  \item \texttt{/} for the main page,
  \item \texttt{/settings} for Wi-Fi configuration,
  \item \texttt{/api/state} for live JSON state,
  \item \texttt{/api/scan} for Wi-Fi scan results,
  \item \texttt{/api/saved} for stored Wi-Fi profiles,
  \item \texttt{/connect} for STA connection request,
  \item \texttt{/forget} for profile reset,
  \item \texttt{/api/light/toggle} for on/off control,
  \item \texttt{/api/light/brightness} for PWM brightness control.
\end{itemize}

\subsubsection*{5) Client-side visualization.}
The AP dashboard includes JavaScript that polls \texttt{/api/state} every 2
seconds. It updates visible values and renders a chart on a
\texttt{canvas}. A recent improvement in the code is the use of dual Y-axes for
temperature and humidity, which avoids the misleading effect of drawing both
series on a single implicit scale. The chart also displays axis labels and
relative sample times.

Two implementation highlights stand out in this module. First, the task uses a
20 ms loop period, which is short enough for responsive HTTP handling without
busy-waiting. Second, the cloud subsystem is not started blindly; instead, the
web task signals \texttt{internetSemaphore} once STA connectivity is confirmed.

\subsection{TinyML Task}
The TinyML task is built around a TensorFlow Lite Micro interpreter and an
8-feature input vector. The implementation sequence is:
\begin{enumerate}
  \item initialize model, resolver, tensor arena, input tensor, and output
  tensor;
  \item receive sensor samples from \texttt{tinyMLQueue};
  \item accumulate ten samples per inference batch;
  \item compute batch-average temperature and humidity;
  \item compute deltas relative to the previous batch;
  \item derive month/hour sine and cosine terms;
  \item normalize all eight features using predefined means and scales;
  \item quantize to the model's input tensor format;
  \item invoke the interpreter and decode output probability.
\end{enumerate}

The resulting feature vector is:
\[
[T_{avg}, H_{avg}, \Delta T, \Delta H, \sin(m), \cos(m), \sin(h), \cos(h)].
\]

This is a thoughtful design. Instead of feeding only raw temperature and
humidity, the model also receives trend information and cyclical time features.
This makes the TinyML task much more meaningful in a real system context.

The task pushes a \texttt{tinyml\_result\_t} structure into
\texttt{tinyResultQueue}, making the result available to the LCD, web, and
CoreIOT tasks. It also prints diagnostic logs containing average values,
temperature/humidity deltas, rain probability, and the final class.

One limitation is that the code currently depends on either valid system time or
fallback month/hour constants. Since the repository contains time utilities but
no active NTP synchronization call in the runtime path, the temporal features
may use fallback values unless time is synchronized elsewhere.

\subsection{CoreIOT Task}
The CoreIOT task is responsible for MQTT connectivity and telemetry publishing.
Its logic is intentionally staged:
\begin{enumerate}
  \item wait on \texttt{internetSemaphore},
  \item configure the MQTT client,
  \item connect to the CoreIOT broker,
  \item subscribe to the RPC request topic,
  \item periodically publish telemetry every 10 seconds.
\end{enumerate}

The telemetry payload includes:
\begin{itemize}
  \item temperature,
  \item humidity,
  \item timestamp (epoch if valid, otherwise tick-based fallback),
  \item rain probability,
  \item rain probability percentage,
  \item categorical rain prediction,
  \item short rain status string,
  \item current LCD forecast mode,
  \item Wi-Fi connection state.
\end{itemize}

This payload is a good design choice because it combines raw measurements with
derived context. The cloud dashboard therefore has richer observability than a
minimal two-field sensor upload.

The CoreIOT task also supports remote control through RPC. It accepts several
method aliases, parses flexible boolean-like parameters, and maps the result to
LCD content mode. In practice, this means a cloud-side control widget can switch
the local LCD between sensor display and TinyML forecast display during STA
operation.

\subsection{Time Utility Module}
The file \texttt{app\_time\_utils.cpp} provides helper routines for:
\begin{itemize}
  \item validating epoch time,
  \item converting epoch time into local date-time fields,
  \item synchronizing with NTP, and
  \item formatting time strings.
\end{itemize}

Although the current runtime path does not actively show a complete NTP setup
sequence, this module is still important because both the TinyML and CoreIOT
parts are time-aware. It creates a clean extension point for future work in
which telemetry timestamps and time-derived inference features are backed by
live synchronized clock values.

\subsection{Implementation Qualities}
Several implementation features are worth highlighting as strengths of the
project:
\begin{itemize}
  \item The system uses \texttt{xQueueOverwrite()} effectively for latest-value
  semantics.
  \item The sensing period is stabilized through \texttt{vTaskDelayUntil()}.
  \item Shared state is reduced and protected by a mutex.
  \item The LCD behavior is policy-driven rather than hard-coded to one screen.
  \item The web interface is not a static page; it actively supports monitoring,
  configuration, and control.
  \item The TinyML task is integrated into the same RTOS pipeline as the other
  system functions.
  \item The CoreIOT task supports both telemetry and remote RPC interaction.
\end{itemize}

Taken together, these implementation choices show that the system is not merely
feature-complete for the assignment; it is architecturally coherent.
